\section{Pseudopotentials}
%
Currently, \textsf{PWDFT.jl} supports a subset of GTH (Goedecker-Teter-Hutter)
pseudopotentials. This type of pseudopotential is analytic and thus is somewhat
easier to program.
%
\textsf{PWDFT.jl} distribution contains several parameters
of GTH pseudopotentials for LDA and GGA functionals.

These pseudopotentials can be written in terms of
local $V^{\mathrm{PS}}_{\mathrm{loc}}$ and
angular momentum $l$ dependent
nonlocal components $\Delta V^{\mathrm{PS}}_{l}$:
\begin{equation}
V_{\mathrm{ene-nuc}}(\mathbf{r}) =
\sum_{I} \left[
V^{\mathrm{PS}}_{\mathrm{loc}}(\mathbf{r}-\mathbf{R}_{I}) +
\sum_{l=0}^{l_{\mathrm{max}}}
V^{\mathrm{PS}}_{l}(\mathbf{r}-\mathrm{R}_{I},\mathbf{r}'-\mathbf{R}_{I})
\right]
\end{equation}

\input{images/PsPot_GTH_struct}

\subsection{Local pseudopotential}

The local pseudopotential for
$I$-th atom, $V^{\mathrm{PS}}_{\mathrm{loc}}(\mathbf{r}-\mathbf{R}_{I})$,
is radially symmetric
function with the following radial form
\begin{equation}
V^{\mathrm{PS}}_{\mathrm{loc}}(r) =
-\frac{Z_{\mathrm{val}}}{r}\mathrm{erf}\left[
\frac{\bar{r}}{\sqrt{2}} \right] +
\exp\left[-\frac{1}{2}\bar{r}^2\right]\left(
C_{1} + C_{2}\bar{r}^2 + C_{3}\bar{r}^4 + C_{4}\bar{r}^6
\right)
\label{eq:V_ps_loc_R}
\end{equation}
with $\bar{r}=r/r_{\mathrm{loc}}$ and $r_{\mathrm{loc}}$, $Z_{\mathrm{val}}$,
$C_{1}$, $C_{2}$, $C_{3}$ and $C_{4}$ are the corresponding pseudopotential
parameters.
In $\mathbf{G}$-space, the GTH local pseudopotential can be written as
\begin{multline}
V^{\mathrm{PS}}_{\mathrm{loc}}(G) = -\frac{4\pi}{\Omega}\frac{Z_{\mathrm{val}}}{G^2}
\exp\left[-\frac{x^2}{2}\right] +
\sqrt{8\pi^3} \frac{r^{3}_{\mathrm{loc}}}{\Omega}\exp\left[-\frac{x^2}{2}\right]\times\\
\left( C_{1} + C_{2}(3 - x^2) + C_{3}(15 - 10x^2 + x^4) + C_{4}(105 - 105x^2 + 21x^4 - x^6) \right)
\label{eq:V_ps_loc_G}
\end{multline}
where $x=G r_{\mathrm{loc}}$.

\subsection{Nonlocal pseudopotential}

\input{images/PsPotNL_struct}

Some notes about \txtinline{prj2beta} variable: prj2beta is currently
using some hardcoded size. It is allocated by the statement:
\begin{juliacode}
prj2beta = Array{Int64}(undef,3,Natoms,4,7)}
\end{juliacode}
The first dimension is the maximum number of projectors per $l$-channel
of GTH pseudopotential.
The second dimension is number of atoms.
The third dimension is the number of angular momentum
The 4th is the m component.
$l$ and $m$ might be combined into one index.

Alternative: from ibeta index and atom index, we can
get iprj, l, and m ?

\begin{juliacode}
NbetaNL = 0
for ia = 1:Natoms
  isp = atm2species[ia]
  psp = Pspots[isp]
  for l = 0:psp.lmax
    for iprj = 1:psp.Nproj_l[l+1]
      for m = -l:l
          NbetaNL = NbetaNL + 1
          prj2beta[iprj,ia,l+1,m+psp.lmax+1] = NbetaNL
      end
    end
  end
end
\end{juliacode}

Alternative:
\begin{juliacode}
prj2beta[1:Natoms][1:Nprojl,l,m]
# l = 1,2,3,4 (s,p,d,f) # 1-based

# l = 1
# m = 1

# l = 2
# m = -1,
\end{juliacode}

This is essentially a way (among several possible ways) to arrange nonlocal
projectors for all atoms.
These projectors will depend on the atomic positions.
Here's how we arrange them:
\begin{juliacode}
for ik = 1:Nkpt
    ibeta = 0
    for ia = 1:Natoms
        isp = atm2species[ia]
        psp = Pspots[isp]
        for l = 0:psp.lmax
        for iprj = 1:psp.Nproj_l[l+1]
        for m = -l:l
            ibeta = ibeta + 1
            idx_gw2g = pw.gvecw.idx_gw2g[ik]
            for igk = 1:Ngw[ik]
                ig = idx_gw2g[igk]
                g[1] = G[1,ig] + k[1,ik]
                g[2] = G[2,ig] + k[2,ik]
                g[3] = G[3,ig] + k[3,ik]
                Gm = norm(g)
                GX = atpos[1,ia]*g[1] + atpos[2,ia]*g[2] + atpos[3,ia]*g[3]
                Sf = cos(GX) - im*sin(GX)
                betaNL[ik][igk,ibeta] =
                (-1.0*im)^l * Ylm_real(l,m,g)*eval_proj_G(psp,l,iprj,Gm,pw.CellVolume)*Sf
            end
        end
        end
        end
    end
end  # kpoints
\end{juliacode}
Note tha the above loop should be done in the way prj2beta is constructed.

Main use case of prj2beta is found in op V Ps nloc:
\begin{juliacode}
for ist in 1:Nstates
    for ia in 1:Natoms
        isp = atm2species[ia]
        psp = Pspots[isp]
        for l in 0:psp.lmax
        for m in -l:l
            for iprj in 1:psp.Nproj_l[l+1], jprj in 1:psp.Nproj_l[l+1]
                #
                ibeta = prj2beta[iprj,ia,l+1,m+psp.lmax+1]
                jbeta = prj2beta[jprj,ia,l+1,m+psp.lmax+1]
                #
                hij = psp.h[l+1,iprj,jprj]
                for ig in 1:Ngw[ik]
                    Vpsi[ig,ist] = Vpsi[ig,ist] + hij*betaNL[ik][ig,ibeta]*betaNL_psi[ist,jbeta]
                end
            end
        end # m
        end # l
    end
end
\end{juliacode}

In this case wen need prj2beta to access ibeta and jbeta indices, for current ia, l, and m.

In some cases, hij is diagonal, so there is no need to use iprj and jprj, for example ONCV.



The nonlocal component of GTH pseudopotential can written in real space as
\begin{equation}
V^{\mathrm{PS}}_{l}(\mathbf{r}-\mathbf{R}_{I},\mathbf{r}'-\mathbf{R}_{I}) =
\sum_{\mu=1}^{N_{l}} \sum_{\nu=1}^{N_{l}} \sum_{m=-l}^{l}
\beta_{\mu lm}(\mathbf{r}-\mathbf{R}_{I})\,
h^{l}_{\mu\nu}\,
\beta^{*}_{\nu lm}(\mathbf{r}'-\mathbf{R}_{I})
\end{equation}
where $\beta_{\mu lm}(\mathbf{r})$ are atomic-centered projector functions
\begin{equation}
\beta_{\mu lm}(\mathbf{r}) =
p^{l}_{\mu}(r) Y_{lm}(\hat{\mathbf{r}})
\label{eq:proj_NL_R}
\end{equation}
%
and $h^{l}_{\mu\nu}$ are the pseudopotential parameters and
$Y_{lm}$ are the spherical harmonics. Number of projectors per angular
momentum $N_{l}$ may take value up to 3 projectors.
%
In $\mathbf{G}$-space, the nonlocal part of GTH pseudopotential can be described by
the following equation.
\begin{equation}
V^{\mathrm{PS}}_{l}(\mathbf{G},\mathbf{G}') =
(-1)^{l} \sum_{\mu}^{3} \sum_{\nu}^{3}\sum_{m=-l}^{l}
\beta_{\mu l m}(\mathbf{G}) h^{l}_{\mu\nu}
\beta^{*}_{\nu l m}(\mathbf{G}')
\end{equation}
with the projector functions
\begin{equation}
\beta_{\mu lm}(\mathbf{G}) = p^{l}_{\mu}(G) Y_{lm}(\hat{\mathbf{G}})
\label{eq:betaNL_G}
\end{equation}
The radial part of projector functions take the following form
\begin{equation}
p^{l}_{\mu}(G) = q^{l}_{\mu}\left(Gr_{l}\right)
\frac{\pi^{5/4}G^{l}\sqrt{ r_{l}^{2l+3}}}{\sqrt{\Omega}}
\exp\left[-\frac{1}{2}G^{2}r^{2}_{l}\right]
\label{eq:proj_NL_G}
\end{equation}
%
For $l=0$, we consider up to $N_{l}=3$ projectors:
\begin{align}
q^{0}_{1}(x) & = 4\sqrt{2} \\
q^{0}_{2}(x) & = 8\sqrt{\frac{2}{15}}(3 - x^2) \\
q^{0}_{3}(x) & = \frac{16}{3}\sqrt{\frac{2}{105}} (15 - 20x^2 + 4x^4)
\end{align}
%
For $l=1$, we consider up to $N_{l}=3$ projectors:
\begin{align}
q^{1}_{1}(x) & = 8 \sqrt{\frac{1}{3}} \\
q^{1}_{2}(x) & = 16 \sqrt{\frac{1}{105}} (5 - x^2) \\
q^{1}_{3}(x) & = 8 \sqrt{\frac{1}{1155}} (35 - 28x^2 + 4x^4)
\end{align}
%
For $l=2$, we consider up to $N_{l}=2$ projectors:
\begin{align}
q^{2}_{1}(x) & = 8\sqrt{\frac{2}{15}} \\
q^{2}_{2}(x) & = \frac{16}{3} \sqrt{\frac{2}{105}}(7 - x^2)
\end{align}
%
For $l=3$, we only consider up to $N_{l}=1$ projector:
\begin{equation}
q^{3}_{1}(x) = 16\sqrt{\frac{1}{105}}
\end{equation}

In the present implementation, we construct the local and nonlocal
components of pseudopotential in the $\mathbf{G}$-space using
their Fourier-transformed expressions
and transformed them back to real space if needed.
We refer the readers to the original
reference \cite{Goedecker1996} and the book \cite{Marx2009}
for more information about GTH pseudopotentials.

Due to the separation of local and non-local components of electrons-nuclei
interaction, the electron-nuclei interaction energy can be written as
\begin{equation}
E_{\mathrm{ele-nuc}} = E^{\mathrm{PS}}_{\mathrm{loc}}
+ E^{\mathrm{PS}}_{\mathrm{nloc}}
\end{equation}
%
The local pseudopotential contribution is
\begin{equation}
E^{\mathrm{PS}}_{\mathrm{loc}} =
\int_{\Omega} \rho(\mathbf{r})\,V^{\mathrm{PS}}_{\mathrm{loc}}(\mathbf{r})\,
\mathrm{d}\mathbf{r}
\end{equation}
%
and the non-local contribution is
\begin{equation}
E^{\mathrm{PS}}_{\mathrm{nloc}} =
\sum_{\mathbf{k}}
\sum_{i}
w_{\mathbf{k}}
f_{i\mathbf{k}}
\int_{\Omega}\,
\psi^{*}_{i\mathbf{k}}(\mathbf{r})
\left[
\sum_{I}\sum_{l=0}^{l_{\mathrm{max}}}
V^{\mathrm{PS}}_{l}(\mathbf{r}-\mathbf{R}_{I},\mathbf{r}'-\mathbf{R}_{I})
\right]
\psi_{i\mathbf{k}}(\mathbf{r})
\,\mathrm{d}\mathbf{r}.
\end{equation}
